Getting the Ball Rolling: Producing a Foundation Text for a Universal Corpus of the World's Languages

### 2014-03-01 docs/acl-2014/paper.tex
Converted all citations to bibtex.

\begin{document}
\maketitle
\begin{abstract}
The study of endangered languages is limited by a lack of data. Existing corpora collections are limited in the range of languages covered, in standardisation, or in machine-readability. This makes the situation even worse for the computational linguist, especially one who would like to take a cross-linguistic or typological approach. We first survey existing efforts to compile cross-linguistic resources, then describe our own approach and give an example application - language clustering. To build the foundation text for a Universal Corpus, we crawled and cleaned texts from several web sources that contain data from a large number languages, and converted them into a standardised form consistent with the guidelines set out by Abney and Bird (2010). The resulting corpus is more easily-accessible and machine-readable than any of the underlying data sources, and represents a significant base corpus for researchers to draw on and add to in the future.
\end{abstract}

@@

\newcite{abneybird2010} posed the grand challenge of building a Universal Corpus, including all of the world's languages, in a consistent structure.

@@

[...]
Currently, multilingual corpora efforts have limited coverage in the number of languages and the number of language families represented. For instance, the OPUS corpus (Tiedemann, 2012) covers over 90 languages, only 1.27\% of the world's languages; The Leipzig Corpora Collection\footnote{corpora.uni-leipzig.de} (???Biemann et al., 2007) contains corpora and dictionaries in 230 languages, 3.24\% of the world's languages. Even corpora that boast of linguistic diversity include only a small number of language families, e.g. the linguistically diverse NTU-Multilingual Corpus (Tan and Bond, 2011) covers only 7 out of 136 language families. Producing a universal corpus requires the linguistic community to merge existing corpora and provide an ubiquitous access interface. To compile the foundation text for the Universal Corpus (uniWaC), we crawled and cleaned web data that contains multilingual texts and merged them with existing corpora collections to form the foundation text for the Universal Corpus.

\subsection{WaCky Corpora}

Baroni and Bernardini (2004) introduced the BootCaT toolkit to bootstrap specialized corpora and terms from the web. Then, Sharoff (2006) built corpora in various languages by issuing queries for random combinations of frequent words to create a balanced corpus not unlike the British National Corpus. Later, Ferraresi (2007) introduced and discussed the notion of compiling Web as Corpus \emph{"properly"} with the inaugural ukWaC using web-crawled data from a list of seed words. Thereafter, the Crúbadán corpora was created for a large number of under-resourced languages (Scannell, 2007). Following which the deWaC (German), itWaC (Italian) and frWaC (French) were compiled (Baroni et al. 2008).

Subsequently, Brunello (2009) stressed the notion of building free corpora from the web using documents released under Creative Commons licenses. To build better quality WaCky corpora, Versley and Panchenko (2012) introduced the notion content-sensitive boilerplate detection in cleaning web corpora.

The initial wave of monolingual WaCky corpora (Baroni et al. 2008) inspired more domain specific WaCky corpora, such as the Korean Web Learner's Corpus (Dickinson et al. 2013) and the Feed Corpus (Minocha et al. 2013). Meanwhile, Web-driven corpora compilations also focused on building large  multilingual resources, e.g. Leipzig Corpora Collection (\texttt{LCC}) (Biemann et al. 2007), the COrpora from the Web (\texttt{COW}) project (Schaefer and Bildhauer, 2012).

@@

\newcite{abneybird2011} describe the data structure in more detail, distinguishing between \textbf{aligned texts} and \textbf{analysed texts}: the former consists of multiple parallel texts, aligned at the document, sentence, or word level; the latter contains more detailed annotations including parts of speech, morphological information, and syntactic relations. In our work, we have strived to maintain a data structure consistent with their recommendations for aligned texts.

However, we disagree that analysed texts should be accommodated in this way. If such a universal corpus is to succeed at all, it must enjoy support from the entire linguistics community, and doing this requires complete theory-neutrality. Although their data structure is fairly lightwight, it is not theory neutral, explicitly encoding support for dependency grammars but not constituency grammars, and positing a specific list of relevant features which should be used to annotate words. Other linguists might wish to use a different set of properties, which could threaten to fragment a universal corpus before it truly gets off the ground. More fundamentally, they assume that the corpus should be segmented into words, but Haspelmath [ref] argues that there is no cross-linguistically valid definition of "word", which would render such an endeavour impossible from the start.

It is not within the scope of this paper to resolve these theoretical concerns. Instead, we suggest that the data structure for a universal corpus should be even more lightweight than Abney and Bird suggest. We agree with their characterisation of aligned texts, but propose an alternative for analysed texts. To motivate the role of parallel texts in a universal corpus, they propose using translations into a high-resource language as a convenient surrogate of meaning. By the same reasoning, we can use the glosses in an IGT to provide a more detailed surrogate of meaning. We propose that, just as we can align texts at the document, sentence, and word levels, we can also align texts at the morpheme level, as recommended by the Leipzig Glossing Rules [ref]. The only difference here is that a text is aligned with descriptions in a metalanguage,\footnote{While we would urge researchers to keep such annotations simple, as usually done in IGTs, there is nothing in principle to stop someone from using a complex metalanguage capable of encoding the full set of properties proposed by Abney and Bird, or even more complicated data structures.} rather than a natural language. The benefits, however, are that we can use the same data structure to represent both aligned and analysed texts, and the representation is simple enough that linguistic controversies cannot be raised. We admit that highly detailed annotations, such as in treebanks and dependency banks, are more awkward to encode, but such resources are unlikely to be available in more than a small handful of languages, and will remain outside the scope of a universal corpus, at least for the foreseeable future.

@@

Although data size matters in general NLP (Banko and Brill, 2001), \emph{universality} is the utmost priority in NLP for the Universal Corpus initiative (aka Human Language Project; Abney and Bird, 2010; 2011). Hence, the uniWaC approach to data collection is based on focused/dedicated crawling unlike previous WaC word/URL list seeding. We manually selected sites that contains texts from a large variety of languages and wrote scripts to crawl and clean specific pages from the different data sources. The crawling and cleaning was done simultaneously.

@@

\begin{minipage}{\columnwidth}
\begin{lstlisting}[language=XML]
<igt>
  <example>
    <line>a. o lesu mai</line>
    <line>2sg return here</line>
    <line>`You return here.'</line>
  </example>
</igit>
\end{lstlisting} 
\centerline{Figure 2: An IGT in ODIN's XML format.}
\end{minipage}
\\ \\
\noindent The IGTs from a web document usually follow the Leipzig Glossing Rules, where the first line is the source language text, the second line contains the word/morphemic equivalence and the third line is an English gloss. From the ODIN XML format, an IGT as of Figure 1 will be represented as in an XML snippet as in Figure 2. Eventually, we need to clean and extract (i) the source text `\emph{o lesu mai}' without the preceding index `\emph{a.}' and (ii) the target language gloss without the quotation marks `\emph{You return here}'.
\begin{minipage}{\columnwidth}
\begin{lstlisting}[language=XML]
<igt>
  <example>
    <line>(69) na-Na-tmi-kwalca-t 
    (Foley 1991)</line>
    <line>3sgA-1sgO-say-rise-PERF
    </line>
    <line>`She woke me up' 
    (by verbal action)</line>
  </example>
</igit>
\end{lstlisting} 
\smallskip
\centerline{Figure 3: Another IGT in ODIN's XML format.}
\end{minipage}
\\ \\
\noindent The primary problem in extracting the source text is lack of consistency of the IGTs from the ODIN data. For instance, Figure 2 uses an alphabet bullet convention, i.e. `\emph{a.}', whereas the IGT in Figure 3 uses a bracketed number indexing convention, i.e. `\emph{(69)}'. In addition, the source line in Figure 3 included the citation to the example, which would be considered as noise to corpus.

Lewis and Xia (2010) noted that it is not uncommon for IGT found in documents to deviant from one uniform convention; when compiling the ODIN data, they attempted a regex based approached and reported 59\% F-score. Similar to their regex approaches for extracting IGT, we cleaned the source line with regexes and kept the ODIN data with two levels of `cleanliness':
\\ \\
\begin{minipage}{\columnwidth}
\begin{itemize}
\item \emph{Cleaner}: Removed (i) all heading and trailing text embedded in square or rounded brackets and (ii) heading double character token ending with bracket or fullstop.
\begin{itemize}
\item[(i)]
\begin{Verbatim}
^(?\s?\w{1,5}\s*[):.]\s*
\end{Verbatim}
\item[(ii)] 
\begin{Verbatim}
[\[\(].{1,}[\]\)]
\end{Verbatim}
\end{itemize}
\item \emph{Cleanest}: Accepts only source lines without any punctuation.
\end{itemize}
\end{minipage}
\\ \\
\noindent The original version of the ODIN data contains XML files for 1275 languages, while the cleaner version of ODIN contains IGTs for 1042 languages and the cleanest version contains IGTs for 402 languages. The drop from 1275 to 1042 languages was largely because {\color{red} XXXX} XML files from the original ODIN data had used language codes that were not in ISO 693-3 and for {\color{red} XXXX} other files, the \texttt{<igt>...</igt>} tags were missing.

@@

Although the redistribution of previous compiled data is not scientifically attractive but redistribution/recompilation of data to improve the usability and quality of the data is worth mentioning. For instance, the SETimes corpus first compiled by Tyers and Alperen (2010) was cleaned and redistributed by Ljubesic and Agic (2013).

>>>>

\begin{document}
\maketitle
\begin{abstract}
The study of endangered languages is limited by a lack of data. Existing corpora collections are limited in the range of languages covered, in standardisation, or in machine-readability. This makes the situation even worse for the computational linguist, especially one who would like to take a cross-linguistic or typological approach. We first survey existing efforts to compile cross-linguistic resources, then describe our own approach and give an example application - language clustering. To build the foundation text for a Universal Corpus, we crawled and cleaned texts from several web sources that contain data from a large number languages, and converted them into a standardised form consistent with the guidelines set out by \newcite{abney2010universal}. The resulting corpus is more easily-accessible and machine-readable than any of the underlying data sources, and represents a significant base corpus for researchers to draw on and add to in the future.
\end{abstract}

@@

\newcite{abney2010universal} posed the grand challenge of building a Universal Corpus, including all of the world's languages, in a consistent structure.

@@

[...]
Currently, multilingual corpora efforts have limited coverage in the number of languages and the number of language families represented. For instance, the OPUS corpus \cite{tiedemann2012opus} covers over 90 languages, only 1.27\% of the world's languages; The Leipzig Corpora Collection\footnote{corpora.uni-leipzig.de} \cite{biemann2007leipzig} contains corpora and dictionaries in 230 languages, 3.24\% of the world's languages. Even corpora that boast of linguistic diversity include only a small number of language families, e.g. the linguistically diverse NTU-Multilingual Corpus \cite{tan2011ntu} covers only 7 out of 136 language families. Producing a universal corpus requires the linguistic community to merge existing corpora and provide an ubiquitous access interface. To compile the foundation text for the Universal Corpus, we crawled and cleaned web data that contains multilingual texts and merged them with existing corpora collections to form the foundation text for the Universal Corpus.

\subsection{Web as Corpus}

\newcite{baroni2004bootcat} introduced the BootCaT toolkit to bootstrap specialized corpora and terms from the web. Then, \newcite{sharoff2006search} built corpora in various languages by issuing queries for random combinations of frequent words to create a balanced corpus not unlike the British National Corpus. Later, \newcite{ferraresi2008ukwac} introduced and discussed the notion of compiling Web as Corpus \emph{"properly"} with the inaugural ukWaC using web-crawled data from a list of seed words. Thereafter, the Crúbadán corpora was created for a large number of under-resourced languages \cite{scannell2007crubadan}. Following which the deWaC (German), itWaC (Italian) and frWaC (French) were compiled \cite{baroni2009wacky}.

Subsequently, \newcite{brunello2009free} stressed the notion of building free corpora from the web using documents released under Creative Commons licenses. To build better quality WaCky corpora, \newcite{versley2012quality} introduced the notion content-sensitive boilerplate detection in cleaning web corpora.

The initial wave of monolingual WaCky corpora \cite{baroni2008wacky} inspired more domain specific WaCky corpora, such as the Korean Web Learner's Corpus \cite{dickinson2010korean} and the Feed Corpus \cite{minocha2013feed}. Meanwhile, Web-driven corpora compilations also focused on building large  multilingual resources, e.g. Leipzig Corpora Collection (\texttt{LCC}) \cite{biemann2007leipzig}, the COrpora from the Web (\texttt{COW}) project \cite{schaefer2012cow}.

@@

\newcite{abney2011data} describe the data structure in more detail, distinguishing between \textbf{aligned texts} and \textbf{analysed texts}: the former consists of multiple parallel texts, aligned at the document, sentence, or word level; the latter contains more detailed annotations including parts of speech, morphological information, and syntactic relations. In our work, we have strived to maintain a data structure consistent with their recommendations for aligned texts.

However, we disagree that analysed texts should be accommodated in this way. If such a universal corpus is to succeed at all, it must enjoy support from the entire linguistics community, and doing this requires complete theory-neutrality. Although their data structure is fairly lightwight, it is not theory neutral, explicitly encoding support for dependency grammars but not constituency grammars, and positing a specific list of relevant features which should be used to annotate words. Other linguists might wish to use a different set of properties, which could threaten to fragment a universal corpus before it truly gets off the ground. More fundamentally, they assume that the corpus should be segmented into words, but \newcite{haspelmath2011segment} argues that there is no cross-linguistically valid definition of "word", which would render such an endeavour impossible from the start.

It is not within the scope of this paper to resolve these theoretical concerns. Instead, we suggest that the data structure for a universal corpus should be even more lightweight than Abney and Bird suggest. We agree with their characterisation of aligned texts, but propose an alternative for analysed texts. To motivate the role of parallel texts in a universal corpus, they propose using translations into a high-resource language as a convenient surrogate of meaning. By the same reasoning, we can use the glosses in an IGT to provide a more detailed surrogate of meaning. We propose that, just as we can align texts at the document, sentence, and word levels, we can also align texts at the morpheme level, as recommended by the Leipzig Glossing Rules.\footnote{http://www.eva.mpg.de/lingua/resources/glossing-rules.php} The only difference here is that a text is aligned with descriptions in a metalanguage,\footnote{While we would urge researchers to keep such annotations simple, as usually done in IGTs, there is nothing in principle to stop someone from using a complex metalanguage capable of encoding the full set of properties proposed by Abney and Bird, or even more complicated data structures.} rather than a natural language. The benefits, however, are that we can use the same data structure to represent both aligned and analysed texts, and the representation is simple enough that linguistic controversies cannot be raised. We admit that highly detailed annotations, such as in treebanks and dependency banks, are more awkward to encode, but such resources are unlikely to be available in more than a small handful of languages, and will remain outside the scope of a universal corpus, at least for the foreseeable future.

@@

Although data size matters in general NLP \cite{banko2001scaling}, \emph{universality} is the utmost priority in NLP for the Universal Corpus initiative (aka Human Language Project; Abney and Bird, 2010; 2011). Hence, the uniWaC approach to data collection is based on focused/dedicated crawling unlike previous WaC word/URL list seeding. We manually selected sites that contains texts from a large variety of languages and wrote scripts to crawl and clean specific pages from the different data sources. The crawling and cleaning was done simultaneously.

@@

\begin{minipage}{\columnwidth}
\begin{lstlisting}[language=XML]
<igt>
  <example>
    <line>a. o lesu mai</line>
    <line>2sg return here</line>
    <line>`You return here.'</line>
  </example>
</igit>
\end{lstlisting} 
\centerline{Figure 2: An IGT in ODIN's XML format.}
\end{minipage}
\\ \\
\noindent The IGTs from a web document usually follow the Leipzig Glossing Rules, where the first line is the source language text, the second line contains the word/morphemic equivalence and the third line is an English gloss. From the ODIN XML format, an IGT as of Figure 1 will be represented as in an XML snippet as in Figure 2. Eventually, we need to clean and extract (i) the source text `\emph{o lesu mai}' without the preceding index `\emph{a.}' and (ii) the target language gloss without the quotation marks `\emph{You return here}'.
\begin{minipage}{\columnwidth}
\begin{lstlisting}[language=XML]
<igt>
  <example>
    <line>(69) na-Na-tmi-kwalca-t 
    (Foley 1991)</line>
    <line>3sgA-1sgO-say-rise-PERF
    </line>
    <line>`She woke me up' 
    (by verbal action)</line>
  </example>
</igit>
\end{lstlisting} 
\smallskip
\centerline{Figure 3: Another IGT in ODIN's XML format.}
\end{minipage}
\\ \\
\noindent The primary problem in extracting the source text is a lack of consistency in the IGTs. In the above examples, the sentence is introduced by a letter or number, which needs to be removed; however, the form of such indexiing elements varies. In addition, the source line in Figure 3 includes a citation for the example, which introduces noise to the corpus.

Lewis and Xia (2010) noted that it is not uncommon for IGTs found in documents to deviate from a single convention; when compiling the ODIN data, they attempted a regex-based approached and reported 59\% F-score. Similar to their regex approaches for extracting IGT, we cleaned the source line with regexes and kept the ODIN data with two levels of `cleanliness':
\\ \\
\begin{minipage}{\columnwidth}
\begin{itemize}
\item \emph{Cleaner}: Removed (i) all heading and trailing text embedded in square or rounded brackets and (ii) heading double character token ending with bracket or fullstop.
\begin{itemize}
\item[(i)]
\begin{Verbatim}
^(?\s?\w{1,5}\s*[):.]\s*
\end{Verbatim}
\item[(ii)] 
\begin{Verbatim}
[\[\(].{1,}[\]\)]
\end{Verbatim}
\end{itemize}
\item \emph{Cleanest}: Accepts only source lines without any punctuation.
\end{itemize}
\end{minipage}
\\ \\
\noindent The original version of the ODIN data contains XML files for 1275 languages, while the cleaner version of ODIN contains IGTs for 1042 languages and the cleanest version contains IGTs for 402 languages. The drop from 1275 to 1042 languages was largely because {\color{red} XXXX} XML files from the original ODIN data had used language codes that were not in ISO 693-3 and for {\color{red} XXXX} other files, the \texttt{<igt>...</igt>} tags were missing.

@@

Although the redistribution of previous compiled data is not scientifically attractive but redistribution/recompilation of data to improve the usability and quality of the data is worth mentioning. For instance, the SETimes corpus first compiled by Tyers and Alperen (2010) was cleaned and redistributed by \newcite{ljubesic2013ner}. [This is not mentioned in the reference...]
